Experimentally informed site-specific substitution models deepen phylogenetic estimates of the divergence of viral lineages

### 2018-02-18 manuscript/manuscript.tex
added in Jesse's comments from slack (no text change)

\topmargin 0.0cm
%\oddsidemargin 2cm
%\textwidth 16cm 
\textheight 21cm
\footskip 1.0cm

@@

\subsection*{Failure to account for site-specificity leads to branch length underestimation.}

@@

@@

\subsection*{empirical Data}

@@

\begin{table}[t!]
\caption{\label{tab:empirical_data}
Performance comparison for the models in~\ref{fig:empirical_trees}. 
All ExpCMs describe the evolution of HA better than the GY94 models, as evaluated by the Akaike information criteria~\citep[$\Delta$AIC,][]{posada2004model}
The mean $\omega$ value from each $K=4$ bins is shown for the models with $\Gamma\omega$ rate variation, otherwise simply the fit, gene-wide $\omega$. 
Each ExpCM has a stringency parameter $>1$.
\skhcomment{GY94+$\Gamma\omega$ has one $\omega$ of 0 because of rounding.}
} 
     \begin{tabular}{cccccccccc}
        \hline
         Model & {\shortstack{Site-specific\\ stationary state}} & $\Gamma\omega$ & $\Delta$AIC & {\shortstack{Log\\ Likelihood}} & {\shortstack{$\omega$\\ (implied dN/dS)}} & {\shortstack{stringency\\ parameter}}\\ \hline
       	ExpCM + $\Gamma\omega$ (H1+H3 avg) & + & + & 0 & -487510 & 0.19,  0.50,  0.90,  1.86 &  1.70\\
	ExpCM (H1+H3 avg) & + & - &  950 & -492270 & 0.15 & 1.78\\
	ExpCM + $\Gamma\omega$ (H1) & + & + & 1306 & -494040  & 0.13 ,  0.44,  0.91,  2.16 & 1.12\\
	ExpCM + $\Gamma\omega$ (H3) & + & + & 1737 & -49620 & 0.09,  0.33,  0.72,  1.77 & 1.28\\
	ExpCM (H1) & + & - & 2556 & -50030 &  0.13 & 1.22\\
	ExpCM (H3) & + & - &  3197 & -50350 & 0.12 & 1.45\\
	GY94 + $\Gamma\omega$ & - & + & 4719 & -51106 & 0.00,  0.03,  0.08,  0.26 & - \\
	GY94 & - & - & 7625 & -52560  & 0.07 & -\\
      \end{tabular}
\end{table}

@@

@@

@@

\begin{suppfig}[H]
\centerline{\includegraphics[width=\textwidth]{figures/prefs_lee}}
\caption{\label{suppfig:prefs_lee}
\textbf{H3 preferences measured by \textit{lee} rescaled with the ExpCM stringency parameter optimized in \ref{fig:tree_lee}A  ($\beta = 1.46$)}
\skhcomment{I need to change the $\beta$ value when the new \texttt{phydms} results finish running.} 
}
\end{suppfig}

\begin{suppfig}[H]
\centerline{\includegraphics[width=\textwidth]{figures/prefs_average}}
\caption{\label{fig:prefs_average}
\textbf{The average of the H1 preferences measured by \cite{doud2016accurate} and the H3 preferences measured by \textit{Lee} rescaled with the ExpCM stringency parameter optimized in \ref{fig:tree_average}A  ($\beta = 1.77$)}}
\skhcomment{I need to change the $\beta$ value when the new \texttt{phydms} results finish running.}
\end{suppfig}

\begin{table}[t!]
\caption{\label{tab:simulation_params}
ExpCM parameters used to simulate sequences in Fig.~\ref{fig:simulation}.}
      \begin{tabular}{ccccc}
        \hline
          Parameter & Value\\ \hline
       	$\beta$ & $1.54$\\
	$\kappa$ & $3.60$\\
	$\omega$ & $0.20$\\
	$\phi_A$, $\phi_C$, $\phi_G$& $0.38$, $0.17$, $0.23$\\
      \end{tabular}
\end{table}

\begin{table}[t!]
\caption{\label{tab:wsn_low_params}
Model parameters used in  Fig.~\ref{fig:decay}.}
      \begin{tabular}{ccccc}
        \hline
          Model & Parameters\\ \hline
          ExpCM & $\beta=1.54196$\\
           & $\kappa=3.47184$\\
           & $\omega=0.219225$\\ 
          YNGKP M0 & $\kappa=2.9984$\\
          & $\omega=0.09076$\\
          YNGKP M5 & $\kappa=2.9984$\\
          & $\omega=0.09076$\\
      \end{tabular}
\end{table}

>>>>

\topmargin 0.0cm
\oddsidemargin 1cm
\textwidth 15cm 
\textheight 21cm
\footskip 1.0cm

@@

\subsection*{Simulations demonstrate how failure to account for site-specificity leads to branch-length underestimation.}

@@

\skhcomment{Change "averaged" to "ExpCM averaged across sites". This will distinguish it from the other types of averages used later. Also clarify in text referring to figure}

@@

\subsection*{Site-specific models estimate longer branches on real data.}

@@

\begin{table}[t!]
\caption{\label{tab:empirical_data}
Performance comparison for the models in~\ref{fig:empirical_trees}. 
All ExpCMs describe the evolution of HA better than the GY94 models, as evaluated by the Akaike information criteria~\citep[$\Delta$AIC,][]{posada2004model}
The mean $\omega$ value from each $K=4$ bins is shown for the models with $\Gamma\omega$ rate variation, otherwise simply the fit, gene-wide $\omega$. 
Each ExpCM has a stringency parameter $>1$.
\skhcomment{GY94+$\Gamma\omega$ has one $\omega$ of 0 because of rounding.}
\skhcomment{check LL values}
} 
     \begin{tabular}{cccccccccc}
        \hline
         Model & {\shortstack{Site-specific\\ stationary state}} & $\Gamma\omega$ & $\Delta$AIC & {\shortstack{Log\\ Likelihood}} & {\shortstack{$\omega$\\ (implied dN/dS)}} & {\shortstack{stringency\\ parameter}}\\ \hline
       	ExpCM + $\Gamma\omega$ (H1+H3 avg) & + & + & 0 & -487510 & 0.19,  0.50,  0.90,  1.86 &  1.70\\
	ExpCM (H1+H3 avg) & + & - &  950 & -492270 & 0.15 & 1.78\\
	ExpCM + $\Gamma\omega$ (H1) & + & + & 1306 & -494040  & 0.13 ,  0.44,  0.91,  2.16 & 1.12\\
	ExpCM + $\Gamma\omega$ (H3) & + & + & 1737 & -49620 & 0.09,  0.33,  0.72,  1.77 & 1.28\\
	ExpCM (H1) & + & - & 2556 & -50030 &  0.13 & 1.22\\
	ExpCM (H3) & + & - &  3197 & -50350 & 0.12 & 1.45\\
	GY94 + $\Gamma\omega$ & - & + & 4719 & -51106 & 0.00,  0.03,  0.08,  0.26 & - \\
	GY94 & - & - & 7625 & -52560  & 0.07 & -\\
      \end{tabular}
\end{table}

\skhcomment{In the beginning of the 4th Results section, you don't need to say we assume site-specific preferences. We aren't assuming anything, we're just analyzing the data. Give a better intro to why we chose HA: there are data sets for two diverged variants (H1 and H3), and this gene has been widely studied using phylogenetics.}

@@

\skhcomment{I know I was pressing on this a lot, but I think that the three trees in Figure 5 turn out not to be useful as they aren't aligned in a helpful way. So I would suggest instead creating something that is a hybrid between the current Supplementary figure 4A and the Figure 5B. It shows the sequence subsets, and then just shows that the stringency parameter goes down for more diverged alignments. I like the Supplementary figure 4A as it shows the trees most clearly. I like the Figure 5B because it gives the stringency parameters but also shows their actual effects in logos, which is more intuitive than just plotting them on a graph.}

@@

\skhcomment{For Figure 6, make a better x-legend. All the models are mutation-selection models. Instead write something like `phylobayes` for the x-legend}

@@

\begin{suppfig}[H]
\centerline{\includegraphics[width=\textwidth]{figures/prefs_lee}}
\caption{\label{suppfig:prefs_lee}
\textbf{H3 preferences measured by \textit{lee} rescaled with the ExpCM stringency parameter optimized in \ref{fig:tree_lee}A  ($\beta = 1.46$)}
 }
\end{suppfig}

\begin{suppfig}[H]
\centerline{\includegraphics[width=\textwidth]{figures/prefs_average}}
\caption{\label{fig:prefs_average}
\textbf{The average of the H1 preferences measured by \cite{doud2016accurate} and the H3 preferences measured by \textit{Lee} rescaled with the ExpCM stringency parameter optimized in \ref{fig:tree_average}A  ($\beta = 1.77$)}}
\end{suppfig}
